---
title: "R Exploratory Data Analysis"
author: DAPT 621 Intro to R
format: html
toc: true
toc-location: left
theme: pulse
editor: visual
---

## Business Scenario

Suppose you are a data consultant recently hired by Company ABC, a mid-size company with 119 employees, to advise on global compensation for analytics positions throughout the company. Currently, the company has three offices: one in Columbus, Ohio, one in Toronto, Canada, and another in Munich, Germany and is shifting from a partially remote set-up to a fully remote one. Company ABC was able to obtain a listing of 607 salaries over 3 years on various technical positions from around the world to use as a benchmark and would like to understand what salaries look like in different locations.

## Data Overview

| Column             | Description                                                                                                                                                                                    |
|-------------|-----------------------------------------------------------|
| work_year          | The year the salary was paid.                                                                                                                                                                  |
| experience_level   | The experience level in the job during the year with the following possible values: EN Entry-level / Junior MI Mid-level / Intermediate SE Senior-level / Expert EX Executive-level / Director |
| employment_type    | The type of employment for the role: PT Part-time FT Full-time CT Contract FL Freelance                                                                                                        |
| job_title          | The role worked in during the year.                                                                                                                                                            |
| salary             | The total gross salary amount paid.                                                                                                                                                            |
| salary_currency    | The currency of the salary paid as an ISO 4217 currency code.                                                                                                                                  |
| salary*in*usd      | The salary in USD (FX rate divided by avg. USD rate for the respective year via fxdata.foorilla.com).                                                                                          |
| employee_residence | Employee's primary country of residence in during the work year as an ISO 3166 country code.                                                                                                   |
| remote_ratio       | The overall amount of work done remotely, possible values are as follows: 0 No remote work (less than 20%) 50 Partially remote 100 Fully remote (more than 80%)                                |
| company_location   | The country of the employer's main office or contracting branch as an ISO 3166 country code.                                                                                                   |
| company_size       | The average number of people that worked for the company during the year: S less than 50 employees (small) M 50 to 250 employees (medium) L more than 250 employees (large)                    |

## Read CSV Data into R

### read.csv

`read.csv` is a base R package (or actually a wrapper around the `read.table` package) specific for reading in comma delimited files. Remember, you can always type `?` before the function to see the function's description and usage.

``` r
?read.csv
```

Let's read in our sample data *ds_salaries.csv* and take a look at the first 5 rows using the `head()` function.

```{r}
filename <- 'ds_salaries.csv'
sal <- read.csv(filename)
```

### head

Notice that 5 rows are returned if no number of rows is specified. Try `head(sal, 10)` to see the first 10 rows.

```{r}
head(sal)
```

### tail

Check the last 5 rows of the data with the `tail()` function.

```{r}
tail(sal)
```

## Explore the data frame

### str

Check the structure of our data object with the `str` function.

```{r}
str(sal)
```

### dim, ncol, nrow

For a quick return of the dimensions, `dim`, `ncol`, and `nrow` can be used on the data frame.

```{r}
dim(sal)
ncol(sal)
nrow(sal)
```

### summary

Use the `summary` function to see summary statistics for each column.

```{r}
summary(sal)
```
